Convolutional Neural Networks and Volcano Plots: Screening and Prediction of Two-Dimensional Single-Atom Catalysts for CO\textsubscript{2} Reduction Reactions

### 2024-03-20 paper/arxiv-submission/main.tex
try to fix "Bibliography not compatible with author-year citations"
fix bibtex
separate content from build details
fix si title page
remove appendix section
global ruff mypy format fix
migrate to windows subsystem for linux
switch to pdf for figures

% Main
\documentclass[a4paper, 12pt, titlepage]{article}

\usepackage{amsmath}
\usepackage{array}
\usepackage{bookmark}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{cleveref}
\usepackage{datetime2}
\usepackage{enumitem}
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{natbib}
\usepackage{nicematrix}
\usepackage{tablefootnote}
\usepackage{textcomp}
\usepackage{siunitx}
\usepackage{xr}
\usepackage[affil-it]{authblk}
\usepackage[english]{babel}
\usepackage[version=4]{mhchem}

\DeclareSIUnit\angstrom{\text {Å}}

\crefname{figure}{Figure}{Figures}
\crefname{table}{Table}{Tables}
\crefname{equation}{Equation}{Equations}
\captionsetup{labelfont=bf}

\externaldocument{supplement}

\begin{document}
\pdfoutput=1

@@

    \date{January 27, 2024}
\maketitle

% Abstract
\newpage
\begin{abstract}
    Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
    yet the diverse combinations of active elements and support materials, the nature of coordination environments,
    elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
    Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models
    to hybrid descriptor based activity volcano plot,
    two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
    including graphitic carbon nitride (g-C\textsubscript{3}N\textsubscript{4}), nitrogen-doped graphene,
    graphene with dual-vacancy, black phosphorous, boron nitride (BN), and C\textsubscript{2}N,
    are screened for efficient CO\textsubscript{2}RR.
    Starting from establishing a correlation map between the adsorption energies of
    intermediates and diverse electronic and elementary descriptors,
    sole singular descriptor lost magic to predict catalytic activity, including d-band center.
    Deep learning method utilizing multi-branch CNN model therefore was employed,
    using 2D electronic density of states (eDOS) as input to predict adsorption energies.
    Hybrid-descriptor enveloping both C- and O-types of CO\textsubscript{2}RR intermediates
    was introduced to construct volcano plots and limiting potential periodic table,
    aiming for intuitive screening of catalyst candidates for efficient CO\textsubscript{2} reduction to CH\textsubscript{4}.
    The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
    To explore the electronic scale principle governing practical engineering catalytic CO\textsubscript{2}RR activity,
    orbitalwise eDOS shifting experiments based on CNN model were employed.
    The study involves examining the adsorption energy and,
    consequently, catalytic activities while varying supported single atoms.
    This work offers a tangible framework to inform both theoretical screening and experimental synthesis,
    thereby paving the way for systematically designing efficient SACs.
\end{abstract}

@@

    In this context, researchers have explored descriptor-based methods
    to circumvent QM calculations and establish direct correlations
    between target properties and electronic factors, commonly referred to as descriptors or features.
    Traditional methods of crafting descriptors, as seen in the d-band model of Hammer and Nørskov \cite{hammer1995electronic},
    necessitate meticulous selection of chemical properties based on chemical knowledge
    and are inherently case-dependent \cite{kajita2017universal}.
    Particularly in the case of SACs, the relationship is intricate,
    rendering manual crafting inadequate \cite{han2021single, thirumalai2018investigating}.
    To streamline this featurization process, we employed CNNs,
    which autonomously derive a hierarchical representation directly from the eDOS
    \cite{tran2015learning, socher2012convolutional, krizhevsky2012imagenet}.
    This approach accurately maps to the activity space, negating the need for manual crafting of descriptors
    and providing a more efficient method to understand and
    establish the structure-mechanism-activity relationship for electrochemical reactions.

@@

% Code/Data Availability, Acknowledgements, Author Contributions and Competing Interests
\appendix
\newpage
    \section{Code and Data Availability}
    Essential source code and data required to replicate the study's results-including VASP input files,
    structure files of SACs, eDOS arrays and the pretrained CNN model-are available on GitHub repository
    at \url{https://github.com/DanielYang59/cnn4dos}, and can be provided by the corresponding authors upon request.
    Due to their substantial size in terabytes, the complete source data (complete VASP output files and such)
    is not currently hosted on publicly available repositories,
    but can be obtained from the corresponding authors upon request.

@@

\newpage
\bibliographystyle{plain}
\bibliography{references}

\end{document}

>>>>

% Main

@@

\date{}
\maketitle

% Abstract
\newpage
\begin{abstract}
    Single-atom catalysts (SACs) have emerged as pivotal frontiers for catalyzing a myriad of chemical reactions,
    yet the diverse combinations of active elements and support materials, the nature of coordination environments,
    elude traditional methodologies in searching optimal SAC systems with superior catalytic performance.
    Herein, by integrating multi-branch Convolutional Neural Network (CNN) analysis models
    to hybrid descriptor based activity volcano plot,
    two-dimensional (2D) SAC system composed of diverse metallic single atoms anchored on six type of 2D supports,
    including graphitic carbon nitride (g-C\textsubscript{3}N\textsubscript{4}), nitrogen-doped graphene,
    graphene with dual-vacancy, black phosphorous, boron nitride (BN), and C\textsubscript{2}N,
    are screened for efficient CO\textsubscript{2}RR.
    Starting from establishing a correlation map between the adsorption energies of
    intermediates and diverse electronic and elementary descriptors,
    sole singular descriptor lost magic to predict catalytic activity, including d-band center.
    Deep learning method utilizing multi-branch CNN model therefore was employed,
    using 2D electronic density of states (eDOS) as input to predict adsorption energies.
    Hybrid-descriptor enveloping both C- and O-types of CO\textsubscript{2}RR intermediates
    was introduced to construct volcano plots and limiting potential periodic table,
    aiming for intuitive screening of catalyst candidates for efficient CO\textsubscript{2} reduction to CH\textsubscript{4}.
    The eDOS occlusion experiments were performed to unravel individual orbital contribution to adsorption energy.
    To explore the electronic scale principle governing practical engineering catalytic CO\textsubscript{2}RR activity,
    orbitalwise eDOS shifting experiments based on CNN model were employed.
    The study involves examining the adsorption energy and,
    consequently, catalytic activities while varying supported single atoms.
    This work offers a tangible framework to guide both theoretical screening and experimental synthesis,
    thereby paving the way for systematically designing efficient SACs.
\end{abstract}

@@

    In this context, researchers have explored descriptor-based methods
    to circumvent QM calculations and establish direct correlations
    between target properties and electronic factors, commonly referred to as descriptors or features.
    Traditional methods of crafting descriptors, as seen in the d-band model of Hammer and Nørskov \cite{hammer1995electronic},
    necessitate meticulous selection of chemical properties based on chemical knowledge
    and are inherently case-dependent \cite{kajita2017universal}.
    Particularly in the case of SACs, the relationship is intricate,
    rendering manual crafting inadequate \cite{han2021single, thirumalai2018investigating}.
    To streamline this featurization process, we employed CNNs,
    which autonomously derive a hierarchical representation directly from the eDOS
    \cite{tran2015learning, socher2012convolutional, NIPS2012_c399862d}.
    This approach accurately maps to the activity space, negating the need for manual crafting of descriptors
    and providing a more efficient method to understand and
    establish the structure-mechanism-activity relationship for electrochemical reactions.

@@

% Code/Data Availability, Acknowledgements, Author Contributions and Competing Interests
\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}

\newpage
    \section{Code and Data Availability}
    Essential source code and data required to replicate the study's results-including VASP input files,
    structure files of SACs, eDOS arrays and the pretrained CNN model-are available on GitHub repository
    at \url{https://github.com/DanielYang59/cnn4dos}, and can be provided by the corresponding authors upon request.
    Due to their substantial size in terabytes, the complete source data (complete VASP output files and such)
    is not currently hosted on publicly available repositories,
    but can be obtained from the corresponding authors upon request.

@@

% End of Main

### 2024-03-20 paper/journal-submission/main.tex
switch to pdf for figures

    % Figure 1: Catalyst Performance Analysis Pipeline
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig1_pipeline.png}
        \caption{\textbf{Catalyst performance analysis pipeline integrating volcano plots and CNN.}
        The figure demonstrates the integrated usage of volcano plots for
        predictive assessment of existing catalysts, and the utility of CNNs
        to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
        eDOS occlusion and shifting experiments are conducted to extract chemical information from CNN model.}
        \label{main_fig1:pipeline}
    \end{figure}

@@

    % Figure 2: CNN for eDOS
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig2_cnn.png}
        \caption{\textbf{CNN for electronic density of states.}
        \textbf{a}. Kendall rank correlation coefficient map illustrating the correlation
        between adsorption energy and other electronic and elementary descriptors.
        Definitions of the notations are provided in
        Supplementary \cref{supp_table15:descriptor_notations}.
        \textbf{b}-\textbf{c}. Architectural of (\textbf{b}) the root and
        (\textbf{c}) each branch of the CNN model.
        The architecture stays consistent across individual branches.
        \textbf{d}-\textbf{m}. Parity plots demonstrating the performance of the
        CNN model across nine adsorbates:
        (\textbf{d}) all adsorbates, (\textbf{e}) CO\textsubscript{2}, (\textbf{f}) COOH,
        (\textbf{g}) CO, (\textbf{h}) CHO, (\textbf{i}) CH\textsubscript{2}O,
        (\textbf{j}) OCH\textsubscript{3}, (\textbf{k}) O, (\textbf{l}) OH, and (\textbf{m}) H.
        The augmentation data are not displayed in these plots, nor included in MAE evaluations.
        Kernel density estimate results are shown along the top and right sides of the main plot,
        representing DFT calculations and CNN model prediction distributions, respectively.}
        \label{main_fig2:cnn_for_eads}
    \end{figure}

@@

    % Figure 3: Limiting Potential Volcano Plot and Periodic Table
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=0.9\textwidth]{main_fig3_volcano.png}
        \caption{\textbf{Limiting potential volcano plot and periodic table.}
        \textbf{a}. Limiting potential volcano plot computed from hybrid-descriptor-based
        linear scaling relations at potential of -0.17 V vs SHE.
        In the plot, the circle, triangle, square, diamond, plus, and star symbols represent
        metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene,
        graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N, respectively.
        Dotted purple lines delineate regions representing rate-determining steps,
        with roman numerals denoting the respective RDS for each region.
        The reaction steps are defined in \cref{main_eq1:co2rr1} to \cref{main_eq8:co2rr8}.
        \textbf{b}. Limiting potential periodic table illustrating the theoretical activity of
        single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.}
        \label{main_fig3:volcano}
    \end{figure}

@@

    % Figure 4: Validation of CNN Predictions
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig4_validation.png}
        \caption{\textbf{Validation of CNN predictions.}
        \textbf{a}-\textbf{c}, The eDOS occlusion experiment results for
        (\textbf{a}) Cr of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{b}) Co of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{c}) Cu of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        eDOS occlusion experiments were conducted on the initial states of CO adsorption process.
        \textbf{d}-\textbf{f}, Real-space wavefunction visualization for
        (\textbf{d}) Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{e}) Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{f}) Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        Wavefunction visualizations were performed on the final states of CO adsorption process.
        The yellow and aqua colors represent areas of positive and negative
        Kohn-Sham orbital coefficients, respectively.
        \textbf{g}-\textbf{j}, Occlusion experiment results for (\textbf{g}) $d_{xy}$ orbital
        and (\textbf{h}) $d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
        COHP analysis for (\textbf{i}) $d_{xy}$ orbital and (\textbf{j}) $d_{xz}$ orbital for
        Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and
        all adsorbate atoms in the final states of CO adsorption process.}
        \label{main_fig4:validation}
    \end{figure}

@@

    % Figure 5: eDOS Shifting Experiments
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig5_shifting.png}
        \caption{\textbf{eDOS shifting experiments.}
        Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
        for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
        The disturbances caused by shifting of (\textbf{a}) entire $d$, (\textbf{b}) $d_{xy}$,
        (\textbf{c}) $d_{yz}$, (\textbf{d}) $d_{z^2}$, (\textbf{e}) $d_{xz}$ and
        (\textbf{f}) $d_{x^2-y^2}$ orbitals are illustrated.
        The shifting step size corresponds to the energy resolution of the eDOS,
        and is 0.005 eV in this study.
        A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
        \label{main_fig5:shifting}
    \end{figure}

@@

    \section{Acknowledgements}
    The QUT eResearch Office provided computational resources and services used in this work.
    This research was undertaken with the assistance of resources and services from
    the National Computational Infrastructure (NCI), supported by the Australian Government.
    It was also supported by resources provided by the Pawsey Supercomputing Research Centre
    with funding from the Australian Government and the Government of Western Australia.
    Insights into deep learning methodologies were contributed by Mr. Yanwei Guan
    from Chongqing University and Mr. Zhipeng He from Assoc. Prof. Chun Ouyang's team at QUT.
    The machine learning and data analysis implementation benefited from
    resources provided by the GitHub, TensorFlow and Stack Overflow communities.
    Dr. Ryky Nelson from RWTH Aachen University offered valuable guidance on
    Crystal Orbital Hamilton Population analysis.

>>>>

    % Figure 1: Catalyst Performance Analysis Pipeline
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig1_pipeline.pdf}
        \caption{\textbf{Catalyst performance analysis pipeline integrating volcano plots and CNN.}
        The figure demonstrates the integrated usage of volcano plots for
        predictive assessment of existing catalysts, and the utility of CNNs
        to predict and modulate adsorption energies \(E_{\text{ads}}\) using eDOS as input.
        eDOS occlusion and shifting experiments are conducted to extract chemical information from CNN model.}
        \label{main_fig1:pipeline}
    \end{figure}

@@

    % Figure 2: CNN for eDOS
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig2_cnn.pdf}
        \caption{\textbf{CNN for electronic density of states.}
        \textbf{a}. Kendall rank correlation coefficient map illustrating the correlation
        between adsorption energy and other electronic and elementary descriptors.
        Definitions of the notations are provided in
        Supplementary \cref{supp_table15:descriptor_notations}.
        \textbf{b}-\textbf{c}. Architectural of (\textbf{b}) the root and
        (\textbf{c}) each branch of the CNN model.
        The architecture stays consistent across individual branches.
        \textbf{d}-\textbf{m}. Parity plots demonstrating the performance of the
        CNN model across nine adsorbates:
        (\textbf{d}) all adsorbates, (\textbf{e}) CO\textsubscript{2}, (\textbf{f}) COOH,
        (\textbf{g}) CO, (\textbf{h}) CHO, (\textbf{i}) CH\textsubscript{2}O,
        (\textbf{j}) OCH\textsubscript{3}, (\textbf{k}) O, (\textbf{l}) OH, and (\textbf{m}) H.
        The augmentation data are not displayed in these plots, nor included in MAE evaluations.
        Kernel density estimate results are shown along the top and right sides of the main plot,
        representing DFT calculations and CNN model prediction distributions, respectively.}
        \label{main_fig2:cnn_for_eads}
    \end{figure}

@@

    % Figure 3: Limiting Potential Volcano Plot and Periodic Table
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=0.9\textwidth]{main_fig3_volcano.pdf}
        \caption{\textbf{Limiting potential volcano plot and periodic table.}
        \textbf{a}. Limiting potential volcano plot computed from hybrid-descriptor-based
        linear scaling relations at potential of -0.17 V vs SHE.
        In the plot, the circle, triangle, square, diamond, plus, and star symbols represent
        metals supported on g-C\textsubscript{3}N\textsubscript{4}, nitrogen-doped graphene,
        graphene with dual-vacancy, black phosphorous, BN, and C\textsubscript{2}N, respectively.
        Dotted purple lines delineate regions representing rate-determining steps,
        with roman numerals denoting the respective RDS for each region.
        The reaction steps are defined in \cref{main_eq1:co2rr1} to \cref{main_eq8:co2rr8}.
        \textbf{b}. Limiting potential periodic table illustrating the theoretical activity of
        single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.}
        \label{main_fig3:volcano}
    \end{figure}

@@

    % Figure 4: Validation of CNN Predictions
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig4_validation.pdf}
        \caption{\textbf{Validation of CNN predictions.}
        \textbf{a}-\textbf{c}, The eDOS occlusion experiment results for
        (\textbf{a}) Cr of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{b}) Co of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{c}) Cu of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        eDOS occlusion experiments were conducted on the initial states of CO adsorption process.
        \textbf{d}-\textbf{f}, Real-space wavefunction visualization for
        (\textbf{d}) Cr $d_{x^2-y^2}$ orbital of Cr@g-C\textsubscript{3}N\textsubscript{4},
        (\textbf{e}) Co $d_{xy}$ orbital of Co@g-C\textsubscript{3}N\textsubscript{4} and
        (\textbf{f}) Cu $d_{xy}$ orbital of Cu@g-C\textsubscript{3}N\textsubscript{4}.
        Wavefunction visualizations were performed on the final states of CO adsorption process.
        The yellow and aqua colors represent areas of positive and negative
        Kohn-Sham orbital coefficients, respectively.
        \textbf{g}-\textbf{j}, Occlusion experiment results for (\textbf{g}) $d_{xy}$ orbital
        and (\textbf{h}) $d_{xz}$ orbital for Co of Co@g-C\textsubscript{3}N\textsubscript{4}.
        COHP analysis for (\textbf{i}) $d_{xy}$ orbital and (\textbf{j}) $d_{xz}$ orbital for
        Co of Co@g-C\textsubscript{3}N\textsubscript{4}, performed between Co and
        all adsorbate atoms in the final states of CO adsorption process.}
        \label{main_fig4:validation}
    \end{figure}

@@

    % Figure 5: eDOS Shifting Experiments
    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\textwidth]{main_fig5_shifting.pdf}
        \caption{\textbf{eDOS shifting experiments.}
        Impact of orbitalwise eDOS shifting on CO adsorption energy, as predicted by the CNN model,
        for single metal atom catalysts supported on g-C\textsubscript{3}N\textsubscript{4}.
        The disturbances caused by shifting of (\textbf{a}) entire $d$, (\textbf{b}) $d_{xy}$,
        (\textbf{c}) $d_{yz}$, (\textbf{d}) $d_{z^2}$, (\textbf{e}) $d_{xz}$ and
        (\textbf{f}) $d_{x^2-y^2}$ orbitals are illustrated.
        The shifting step size corresponds to the energy resolution of the eDOS,
        and is 0.005 eV in this study.
        A positive eDOS shift indicates a shift towards higher energy levels, and vice versa.}
        \label{main_fig5:shifting}
    \end{figure}

@@

    \section{Acknowledgements}
    We acknowledge funding support from the Australian Research Council
    (FT160100281, FT180100387, DP200103568, DP230101625).
    The QUT eResearch Office provided computational resources and services used in this work.
    This research was undertaken with the assistance of resources and services from
    the National Computational Infrastructure (NCI), supported by the Australian Government.
    It was also supported by resources provided by the Pawsey Supercomputing Research Centre
    with funding from the Australian Government and the Government of Western Australia.
    Insights into deep learning methodologies were contributed by Mr. Yanwei Guan
    from Chongqing University and Mr. Zhipeng He from Assoc. Prof. Chun Ouyang's team at QUT.
    The machine learning and data analysis implementation benefited from
    resources provided by the GitHub, TensorFlow and Stack Overflow communities.
    Dr. Ryky Nelson from RWTH Aachen University offered valuable guidance on
    Crystal Orbital Hamilton Population analysis.
